---
title: "FARS Functions"
author: "Katalin Virag"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE, message = FALSE}
library(FARS)
g <- system.file("extdata", package = "FARS")
knitr::opts_knit$set(root.dir = g)
```

These functions can be used to analyse data from the US National Highway Traffic Safety Administration's Fatality Analysis Reporting System, which is a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes.

## Functions included in the package

* `fars_read()` to read data.
* `fars_summarize_years()` to summarize data.
* `fars_map_state()` to create a map of the events for given state and year.

## Data

This package includes three datasets regarding fatal injuries suffered in motor vehicle traffic crashes in 2013, 2014, and 2015.

### Read in data with `fars_read()`

`fars_read()` allows you to read a comma-separated csv file into a tibble.

The *accident_2015.csv.bz2* file contains the features of events occured in 2015. 

```{r}
fname <- "accident_2015.csv.bz2"
fpath <- system.file("extdata", fname, package = "FARS")
d <- fars_read(fpath)
```

## Analyse data

### Summarize with `fars_summarize_years()`

`fars_summarize_years()` returns with the number of observations per month in the requested year(s).
Such summary statistics for years 2013 and 2015 can be obtained as follows:

```{r}
fars_summarize_years(c(2013, 2015))
```

### Map of accidents with `fars_map_state()`

`fars_map_state()` creates a map of events for given state and year.

The map of the accidents occured in 2015 in *state 1* can be obtained as follows:

```{r}
fars_map_state(1, 2015)
```
